---
title: "Project"
author: "Brandon Sucher"
date: "2023-04-07"
output: html_document
---

Read in data
```{r}
features_test <- read.csv("dengue_features_test.csv")
features_train <- read.csv("dengue_features_train.csv")
labels_train <- read.csv("dengue_labels_train.csv")
```

Test some automated EDA
```{r}
pacman::p_load(tidyverse, DataExplorer, tableone)
all_train <- left_join(features_train, labels_train, by = c("city", "year", "weekofyear"))

all_train %>% create_report(output_file = "EDA",
                            report_title = "EDA Report - Dengue Data",
                            y = "total_cases")

CreateTableOne(vars = colnames(select(all_train, -c("city", "year", "weekofyear",
                                                    "week_start_date"))),
               strata = c("city"),
               data = all_train)
```

Notes:
- ndvi_ne has 13% missingness but 191/194 missing values in that variable are from San Juan, PR. Which makes that 20.4% missingness of this variable for SJ.
- There's a few outliers of total_cases. We'll have to decide how to treat these.
- But station precipitation kind of follows the same trend...
- reanalysis_tdr_k also seems like to have some interesting diagnostics...
- total_cases not strongly correlated with any one variable
- the higher number of total_cases were really only in 2 different years
- Big difference in precipiation amount between SJ and IQ
- Also a huge difference between total_cases average between the two cities
- Station precipitation is also much higher in IQ than SJ but total_cases much lower in IQ than SJ

