---
title: "Project"
author: "Brandon Sucher"
date: "2023-04-07"
output: html_document
---

Read in data
```{r}
features_test <- read.csv("dengue_features_test.csv")
features_train <- read.csv("dengue_features_train.csv")
labels_train <- read.csv("dengue_labels_train.csv")
```

Test some automated EDA
```{r}
pacman::p_load(tidyverse, DataExplorer, tableone)
all_train <- left_join(features_train, labels_train, by = c("city", "year", "weekofyear"))

all_train %>% create_report(output_file = "EDA",
                            report_title = "EDA Report - Dengue Data",
                            y = "total_cases")

CreateTableOne(vars = colnames(select(all_train, -c("city", "year", "weekofyear",
                                                    "week_start_date"))),
               strata = c("city"),
               data = all_train)
```

Notes:
- ndvi_ne has 13% missingness but 191/194 missing values in that variable are from San Juan, PR. Which makes that 20.4% missingness of this variable for SJ.
- There's a few outliers of total_cases. We'll have to decide how to treat these.
- But station precipitation kind of follows the same trend...
- reanalysis_tdr_k also seems like to have some interesting diagnostics...
- total_cases not strongly correlated with any one variable
- the higher number of total_cases were really only in 2 different years
- Big difference in precipitation amount between SJ and IQ
- Also a huge difference between total_cases average between the two cities
- Station precipitation is also much higher in IQ than SJ but total_cases much lower in IQ than SJ

```{r}
ggplot(all_train, aes(x = station_precip_mm, y = total_cases, color = city)) +
  geom_point() +
  labs(x = "Specific Humidity (g/kg)", y = "Density", color = "City") +
  theme_bw()
```

```{r}
veg_data <- all_train[,c(1:3, 5:8, 25)]
colnames(veg_data) <- c("City", "year", "weekofyear", "Northeast", "Northwest",
                        "Southeast", "Southwest", "total_cases")
veg_data <- pivot_longer(veg_data, cols = c("Northeast", "Northwest", 
                                            "Southeast", "Southwest"),
             names_to = "Direction")
veg_data <- veg_data[veg_data$weekofyear != 53,]
veg_data$Date <- as.Date(paste(veg_data$year, veg_data$weekofyear, 1, sep="-"),
                         "%Y-%U-%u")
# par(mfrow = c(2,1))
ggplot(veg_data, aes(x = Date, y = value, col = City)) +
  geom_line() +
  geom_vline(data = veg_data[which(veg_data$total_cases > 50 & veg_data$City == "iq"),],
             aes(xintercept = Date), linetype = 4, color = "black", alpha = 0.5) +
  geom_vline(data = veg_data[which(veg_data$total_cases > 100 & veg_data$City == "sj"),],
             aes(xintercept = Date), linetype = 4, color = "yellow3", alpha = 0.5) +
  # geom_point(aes(y = total_cases)) +
  facet_grid(Direction ~ ., scales = "free") +
  scale_color_brewer(palette = "Dark2",
                     labels = c("Iquitos", "San Juan")) +
  labs(y = "Normalized Difference Vegetation Index (NDVI)",
       title = "Vegetation Index Over Time") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
# ggplot(veg_data, aes(x = week, y = total_cases, col = city)) +
#   geom_line()
```

```{r}
all_train <- all_train[all_train$weekofyear != 53,]
all_train$week <- as.Date(paste(all_train$year, all_train$weekofyear, 1, sep="-"),
                         "%Y-%U-%u")
ggplot(all_train[all_train$city == "sj",], aes(x = week, y = total_cases)) +
  geom_line(col = "#d95f02") +
  theme_bw() +
  labs(x = "Date", y = "Total Cases",
       title = "Total Number of Weekly Cases Over Time in San Juan, Puerto Rico") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
veg_data <- all_train[,c(1:3, 5:8, 25)]
veg_data <- veg_data[veg_data$weekofyear != 53,]
veg_data$Date <- as.Date(paste(veg_data$year, veg_data$weekofyear, 1, sep="-"),
                         "%Y-%U-%u")
for (i in 1:nrow(veg_data)) {
  veg_data$sum_veg[i] <- sum(veg_data$ndvi_ne[i], veg_data$ndvi_nw[i], 
                        veg_data$ndvi_se[i], veg_data$ndvi_sw[i])
}

ggplot(veg_data, aes(x = Date, y = sum_veg, col = city)) +
  geom_line() +
  scale_color_brewer(palette = "Dark2",
                     labels = c("Iquitos", "San Juan")) +
  geom_vline(data = veg_data[which(veg_data$total_cases > 50 & veg_data$city == "iq"),],
             aes(xintercept = Date), linetype = 4, color = "black", alpha = 0.5) +
  geom_vline(data = veg_data[which(veg_data$total_cases > 100 & veg_data$city == "sj"),],
             aes(xintercept = Date), linetype = 4, color = "yellow3", alpha = 0.5) +
  theme_bw() +
  labs(y = "Sum of NDVI Values", color = "City",
       title = "Summed Vegetation Index Over Time") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
for (i in 1:nrow(veg_data)) {
  veg_data$avg_veg[i] <- sum(veg_data$ndvi_ne[i], veg_data$ndvi_nw[i], 
                        veg_data$ndvi_se[i], veg_data$ndvi_sw[i])/4
}

ggplot(veg_data, aes(x = Date, y = avg_veg, col = city)) +
  geom_line() +
  geom_vline(data = veg_data[which(veg_data$total_cases > 50 & veg_data$city == "iq"),],
             aes(xintercept = Date), linetype = 4, color = "black", alpha = 0.2) +
  geom_vline(data = veg_data[which(veg_data$total_cases > 100 & veg_data$city == "sj"),],
             aes(xintercept = Date), linetype = 4, color = "yellow", alpha = 0.2)
```

